---
title: "Linear Model w/ Prediction"
author: "Tyler Hamblin"
date: "19 October 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
data <- read.csv(file = "https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)
```

## Model 1 - SIMS~ARM
### Scatterplot

```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs. ARM")
```

The scatter plot above shows a positive linear progression as the data for ARM increases from left to right, the intercept on the SIMS axis increases from top to bottom. 


### Numerical Results

```{r}
cor(SIMS~ARM,data=data)
```

These numerical results confirm the positive relationship between SIMS explained by ARM that the graphical data above suggested. 


### Inferential Results 

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```




### Create Target Point

```{r}
target <- data.frame(ARM=88, GRIP=94)
```




### Predict Target Point for Model 1 

```{r}
predict(model.1, target, interval = c("prediction"))
```



### Scatterplot for Model 1 w/ Model Fit 


```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs. ARM") + geom_smooth(method=lm)
```



## Model 2 - SIMS~GRIP
### Scatterplot


```{r}
basicGRIP <- ggplot(data,aes(y=SIMS,x=GRIP))
```

```{r}
basicGRIP + geom_point() + ggtitle("SIMS vs. GRIP")
```

The scatter plot suggests the grip strength that a worker would need for the job. It shows that most of the population requires a grip strength of about 80 to 130. 



### Numerical Results

```{r}
cor(SIMS~GRIP,data=data)
```

Again the numerical data confirms the relationship of the above graph and is only lower than Model 1's numerical results by about 0.05. 


### Inferential Results 

```{r}
model.1 <- lm(SIMS~GRIP,data=data)
summary.lm(model.1)
```




















