---
title: "Linear Model w/ Prediction"
author: "Tyler Hamblin"
date: "19 October 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
data <- read.csv(file = "https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)
```

## Model 1 - SIMS~ARM
### Scatterplot

```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs. ARM")
```

The scatter plot above shows a positive linear progression as the data for ARM increases from left to right, the intercept on the SIMS axis increases from top to bottom. 


### Numerical Results

```{r}
cor(SIMS~ARM,data=data)
```

These numerical results confirm the positive relationship between SIMS explained by ARM that the graphical data above suggested. 


### Inferential Results 

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```




### Create Target Point

```{r}
target <- data.frame(ARM=88, GRIP=94)
```




### Predict Target Point for Model 1 

```{r}
predict(model.1, target, interval = c("prediction"))
```



### Scatterplot for Model 1 w/ Model Fit 


```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs. ARM") + geom_smooth(method=lm)
```



## Model 2 - SIMS~GRIP
### Scatterplot


```{r}
basicGRIP <- ggplot(data,aes(y=SIMS,x=GRIP))
```

```{r}
basicGRIP + geom_point() + ggtitle("SIMS vs. GRIP")
```

The scatter plot suggests the grip strength that a worker would need for the job. It shows that most of the population requires a grip strength of about 80 to 130. 



### Numerical Results

```{r}
cor(SIMS~GRIP,data=data)
```

Again the numerical data confirms the spread of the above graph and is only lower than Model 1's numerical results by about 0.05. This suggests that there is slightly worse positive linear relationship for SIMS defined by GRIP. 


### Inferential Results 

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```




### Predict Target Point for Model 2 

```{r}
predict(model.2, target, interval = c("prediction"))
```



### Scatterplot for Model 2 w/ Model Fit 


```{r}
basicGRIP <- ggplot(data,aes(y=SIMS,x=GRIP))
```

```{r}
basicGRIP + geom_point() + ggtitle("SIMS vs. GRIP") + geom_smooth(method=lm)
```

When comparing the SIMS vs. ARMS and the SIMS vs. GRIP scatter plots, we can tell that the Model 1 has more points scattered towards the far ends of the target point line while Model 2 has a tighter grouping towards the middle of the graph. It may appear that Model 2 has a better fit for the target but according to the numerical data that interprets the plots, SIMS is better explained by the variable ARMS. 


## Model 3 - SIMS~ARM+GRIP
### Scatterplot

```{r}
basicARMGRIP <- ggplot(data,aes(y=SIMS,x=ARM+GRIP))
```

```{r}
basicARMGRIP + geom_point() + ggtitle("SIMS vs. ARM+GRIP")
```

Again for the last explanation of SIMS now defined by the variables ARM AND GRIP, we see a positive linear progression to the right. The majority of data falls between the 150 - 250 X-value range. 


### Numerical Results 
```{r}
cor(SIMS~ARM+GRIP, data=data)
```

This value is significantly higher than the previous two, showing that the best way to influence SIMS is by considering the ARM and GRIP variables at the same time. 


### Inferential Results 
```{r}
model.3 <- lm(SIMS~ARM+GRIP,data=data)
summary.lm(model.3)
```













