---
title: "Linear Model w/ Prediction"
author: "Tyler Hamblin"
date: "19 October 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
data <- read.csv(file = "https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)
```

## Model 1 - SIMS~ARM
### Scatterplot

```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs. ARM")
```

The scatter plot above shows a positive linear progression as the data for ARM increases from left to right, the intercept on the SIMS axis increases from top to bottom. 


### Numerical Results

```{r}
cor(SIMS~ARM,data=data)
```

These numerical results confirm the positive relationship between SIMS explained by ARM that the graphical data above suggested. 


### Inferential Results 

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```




### Create Target Point

```{r}
target <- data.frame(ARM=88, GRIP=94)
```




### Predict Target Point for Model 1 

```{r}
predict(model.1, target, interval = c("prediction"))
```



### Scatterplot for Model 1 w/ Model Fit 


```{r}
basicNN <- ggplot(data,aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + ggtitle("SIMS vs. ARM") + geom_smooth(method=lm)
```



## Model 2 - SIMS~GRIP
### Scatterplot


```{r}
basicGRIP <- ggplot(data,aes(y=SIMS,x=GRIP))
```

```{r}
basicGRIP + geom_point() + ggtitle("SIMS vs. GRIP")
```

The scatter plot suggests the grip strength that a worker would need for the job. It shows that most of the population requires a grip strength of about 80 to 130. 



### Numerical Results

```{r}
cor(SIMS~GRIP,data=data)
```

Again the numerical data confirms the spread of the above graph and is only lower than Model 1's numerical results by about 0.05. This suggests that there is slightly worse positive linear relationship for SIMS defined by GRIP. 


### Inferential Results 

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```




### Predict Target Point for Model 2 

```{r}
predict(model.2, target, interval = c("prediction"))
```



### Scatterplot for Model 2 w/ Model Fit 


```{r}
basicGRIP <- ggplot(data,aes(y=SIMS,x=GRIP))
```

```{r}
basicGRIP + geom_point() + ggtitle("SIMS vs. GRIP") + geom_smooth(method=lm)
```

When comparing the SIMS vs. ARMS and the SIMS vs. GRIP scatter plots, we can tell that the Model 1 has more points scattered towards the far ends of the target point line while Model 2 has a tighter grouping towards the middle of the graph. It may appear that Model 2 has a better fit for the target but according to the numerical data that interprets the plots, SIMS is better explained by the variable ARMS. 


## Model 3 - SIMS~ARM+GRIP
### Scatterplot

```{r}
basicARMGRIP <- ggplot(data,aes(y=SIMS,x=ARM+GRIP))
```

```{r}
basicARMGRIP + geom_point() + ggtitle("SIMS vs. ARM+GRIP")
```

Again for the last explanation of SIMS now defined by the variables ARM AND GRIP, we see a positive linear progression to the right. The majority of data falls between the 150 - 250 X-value range. 


### Numerical Results 
```{r}
cor(SIMS~ARM+GRIP, data=data)
```

This value is significantly higher than the previous two, showing that the best way to influence SIMS is by considering the ARM and GRIP variables at the same time. 


### Inferential Results 
```{r}
model.3 <- lm(SIMS~ARM+GRIP,data=data)
summary.lm(model.3)
```



### Predict Target Point for Model 2 

```{r}
predict(model.3, target, interval = c("prediction"))
```

### Scatterplot for Model 3 w/ Model Fit 


```{r}
basicGRIP <- ggplot(data,aes(y=SIMS,x=ARM+GRIP))
```

```{r}
basicGRIP + geom_point() + ggtitle("SIMS vs. GRIP") + geom_smooth(method=lm)
```

Even visually, comparing Model 3 with the line of best fit, we can see that this graph has the tightest grouping of points closest to the blue line. There are still a few outliers, especially when approaching the 300 range, but towards the middle the points are extremely close, sometimes overlapping. With the numerical and graphical results we can confidently say that ARM+GRIP is the most influential factor defining SIMS. 


## Comparing Models Using Anova
### Model 1 vs Model 3

```{r}
anova(model.1,model.3)
```
The P-value for the comparison of Model 1 and Model 3 is 0.000004994, a very insignificant value. This gives little random chance that 3 just happens to be better 1. 

### Model 2 vs Model 3

```{r}
anova(model.2,model.3)
```


The p-value for the comparison between Models 2 and 3 is 0.000000001495. Again a very small number leaves less room for the error of an accidental statistical relationship affecting SIMS. Model 3 can be considered the best option as it beats both 1 and 2 in it's accuracy to the line of best fit. 


### Model 1 vs Model 2 + Conclusion

```{r}
anova(model.1,model.2)
```

Model 1 appears to be slightly better than model 2 so we know ranked from best to worst the Models go 3, 1, and 2. This concludes the statistical analysis as to what is the best option from this data set to define SIMS. After repeating the same structure of graphical, numerical, and inferential commands, only changing the inputs to the code, we are better able to side by side compare each variable. We also incorporated the use of prediction and "geom_smooth(method=lm)" commands to allow us to see how ARM, GRIP, and ARM+GRIP compare to a line that perfectly explains SIMS based on the value of the X-axis. 



